Predicting Remaining Useful Life of NASA Aircraft Engines 

### 2025-02-15 index.qmd
Update Intro and references

Slides: [slides.html](slides.html){target="_blank"} ( Go to `slides.qmd`
to edit)

[Literature](articleSummaries.html)

@@

As the COVID-19 pandemic demonstrated, the ability to predict the
survival of a patient based on their symptoms and medical history is a
major benefit as society deals with health crisis like pandemics. Those
predictions would assist medical practitioners in triage and allow
researchers the ability to generate predictions on the individuals who
are at the greatest risk of losing their lives due to an illness. While
the medical world is the most obvious use of survival predictions, the
same methods can be applied to the predicting bank failure, the
remaining useful life of machines, machinery maintenance timelines, and
insurance likelihood of payouts. There are various survival methods that
can be examined, but this paper will focus on Cox proportional hazards
model. (*don't like)*

Cox proportional hazards (CPH) model is a statistical regression model
for survival modeling that specializes in time-to-event predictions
using using single or multiple covariates and the effects they have on
survival. Traditional regression models struggle to capture the
time-to-event nature of data and treat the time variable as a fixed
value instead of continuous variable. CPH models are able to overcome
those obstacles and are capable of dealing with survival data that
contains censored (events that have not occurred yet in the study
period) data as well. The models also have the advantage of being
semi-parametric which allows more flexibility as compared to other
models.

@@

*Intro guidelines:* The introduction should:

-   Develop a storyline that captures attention and maintains interest.

-   Your audience is your peers

-   Clearly state the problem or question you're addressing.

@@

-   Introduce why it is relevant needs.

-   Provide an overview of your approach.

Example of writing including citing references:

*This is an introduction to ..... regression, which is a non-parametric
estimator that estimates the conditional expectation of two variables
which is random. The goal of a kernel regression is to discover the
non-linear relationship between two random variables. To discover the
non-linear relationship, kernel estimator or kernel smoothing is the
main method to estimate the curve for non-parametric statistics. In
kernel estimator, weight function is known as kernel function
[@efr2008]. Cite this paper [@bro2014principal]. The goat study*
[@ziadi2023]*. The insurance[@zapletal2021]*

*This is my work and I want to add more work...*

>>>>

Slides: [slides.html](slides.html){target="_blank"}

Literature Comments: [Literature.html](articleSummaries.html)

@@

As the COVID-19 pandemic demonstrated, the ability to predict the
survival of a patient based on their symptoms and medical history is a
major benefit as society deals with health crisis like pandemics. Those
predictions would assist medical practitioners in triage and allow
researchers the ability to generate predictions on the individuals who
are at the greatest risk of losing their lives due to an illness. While
the medical world is the most obvious use of survival predictions, the
same methods can be applied to the predicting bank failure, the
remaining useful life of machines, machinery maintenance timelines, and
insurance likelihood of payouts. There are various survival methods that
can be examined, including Cox proportional hazards model which will be
examined in this paper.

Cox proportional hazards (CPH) model is a statistical regression model
for survival modeling that specializes in time-to-event predictions
using using single or multiple covariates and the effects they have on
survival. Traditional regression models struggle to capture the
time-to-event nature of data and treat the time variable as a fixed
value instead of continuous variable. CPH models are able to overcome
those obstacles and are capable of dealing with survival data that
contains censored data as well. Censored data is when the information
about an individual in a study is only known for a certain period of
time, such as the length of the study [@klein2005, 1]. The models also
have the advantage of being semi-parametric which allows more
flexibility as compared to other models.

As mentioned, CPH is used in various fields including the health
industry, but the usage of the model does not come without challenges. A
COVID-19 study done using data from a Pakistani hospital noted that the
model was limited due to the recovery possibility of COVID-19 patients
as CPH assumes as time goes on, the survival probability will approach
zero with no survivors. Though CPH was not the best model for this
study, the results of the model were that individuals who were
asymptomatic and young had a higher chance of surviving the virus
[@asghar2024]. In 2024, a study was done in the US on the relationship
between chest pain occurring and mortality of the person experiencing
chest pain using CPH. A response to that study was written that provided
the limitations of CPH with the biggest concern being the assumption of
constant hazards ratio that CPH relies on for the model. In the heart
study, the effect of Triglyceride-Glucose was assumed to be a constant,
but is actually a value that fluctuates over time along with an
individual's body weight and blood glucose levels. As such, the model
would struggle to predict correctly the effect of Triglyceride-Glucose
given the violation of the constant hazards ratio [@jiang2024].

An additional limitation of CPH comes from the data in the covariaties
selected and how can they can become biased. Two studies done using UK
health data demonstrated how bias can impact the values for the
covariates when relying on patient submitted information. The first was
a study on the connection between salt intake and the development of
anxiety or depression after 14.5 years [@wang2025] and the second was a
study on the association between postpartum depression and the
development of two or more chronic diseases [@zhang2025]. Both of those
studies relied on patient submitted health data and survey responses
that were supplemented with death records and other accessible health
records thus creating a possible bias in their data which could
negatively influence their model as their covariates may not have
accurately reflected the true data. These articles provide examples of
the care that needs to be taken when working with CPH models.

While health care is the dominate field in which CPH is used, a study in
2021 provides an example of using CPH in insurance fields. An insurance
company in the Czech Republic provides policies for certain illness
events until the event occurs when the insured will receive money and
their contract is complete. To assist the company with predicting if an
individual will have one of the illness events occur, a model using CPH
was generated using gender, age, and region as the
covariates[@zapletal2021]. A study in 2023 focused on the length of
productive life of female Floridian goats to determine the factors that
would impact the longevity of the goats. CPH was used to determine that
the age of first kidding and the specific herd impacts the length of
productive life of the goats [@ziadi2023]. Both of these studies provide
examples of the reaches of CPH models outside the health industry.

Continuing with using CPH outside of health data, we will study the
remaining useful life of a NASA aircraft engine using data provided by
NASA for the study on propagation modeling [@saxena2008]. From the
literature review, we have noted that we will need to be concerned with
the constant hazards ratio within our data and determine how we will
censor our data.

@@

@@
